---
title: "ESM 204 Assignment 4"
author: "Naomi Raal, Virgina Pan, and Cam Audras"
date: "5/22/2022"
output: 
  html_document:
    theme: yeti
    toc: yes
    code_folding: hide # hide code
---

```{r setup, include=FALSE, echo=FALSE, mesage=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

```{r}
damage <- read_csv(here("damages.csv"))
warming <- read_csv(here("warming.csv"))
```

### **1. Damages vs. Change in Temperature**
Using damages.csv, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature. Estimate a model with a zero intercept because damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data.

```{r}
#create linear regression 

damage$warming_squared <- damage$warming^2
quad_damage_model <- lm(damages ~ warming + warming_squared + 0, data = damage)

# define function 
quad_function <- function(warming){
  harm <- quad_damage_model$coefficients[1]*warming + quad_damage_model$coefficients[2]*warming^2
  return(harm)
}
```


```{r}
# visualize function and scatterplot graph
ggplot(quad_damage_model, aes(warming, damages)) + geom_smooth(method = "lm", formula = y ~ x + I(x^2), color = "brown") + geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damage Function", x = "Warming in degrees C", y = "Annual Damages in USD")
```


### **2. Predicted Damages per Year**
Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate and the pulse scenario. Make four plots: (1) damages over time without the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that arises from the pulse, and (4) the difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage).

(1) damages in each year under baseline climate scenario
```{r}
warming$damage_baseline <- quad_function(warming$warming_baseline)
```
```{r}
# visualize function and scatterplot graph
ggplot(data = warming, aes(year, damage_baseline)) + 
  geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damages under baseline", x = "time", y = "Annual Damages in USD")
```

(2) damages in each year under baseline + pulse climate scenario
```{r}
warming$damage_pulse <- quad_function(warming$warming_pulse)
```

```{r}
# visualize function and scatterplot graph
ggplot(data = warming, aes(year, damage_pulse)) + 
  geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damages under baseline with pulse", x = "time", y = "Annual Damages in USD")
```

(3) damages in each year under pulse alone climate scenario
```{r}
warming$damage_pulse_no_baseline <- warming$damage_pulse-warming$damage_baseline
```

```{r}
# visualize function and scatterplot graph
ggplot(data = warming, aes(year, damage_pulse_no_baseline)) + 
  geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damages under pulse alone (no baseline)", x = "time", y = "Annual Damages in USD")
```

(4) difference in damages over time per ton of CO2
```{r}
warming$unit_damage <- (warming$damage_pulse_no_baseline)/(35*10^9)
```

```{r}
# visualize function and scatterplot graph
ggplot(data = warming, aes(year, unit_damage)) + 
  geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damages from pulse per ton of CO2", x = "time", y = "Annual Damages in USD per ton CO2")
```



### **3. Investigating the Relationship between Discount Rate and SCC**
The SCC is the present discounted value of the stream of future damages caused by one
additional ton of CO2. The Obama Administration used a discount rate of 3% to discount
damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot
of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates.
Explain the intuition for how the discount rate affects the SCC.

```{r}
# define SCC calculation function 
SCC_calc <- function(principal, r, t){
  r = r/100
  SCC <- principal/((1+r)^t)
  return(SCC)
}
```

```{r}
names(warming)[names(warming) == '...1'] <- 'time'
discount_rates_SCC <- warming[, c("time", "unit_damage")]
discount_rates_SCC$r_1 <- SCC_calc(warming$unit_damage,1,warming$time)
total_SCC_1 <- sum(warming$SCC_1)
```



### **4. Using the Ramsey Rule to find SCC**
The National Academies of Sciences, Engineering, and Medicine advised the government in a
2017 report to use the Ramsey Rule when discounting within the SCC calculation:
r = ρ + ηg
Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from
above.


### **5. Comparing Potential Policy Outcomes**
Now suppose there are two possible climate policies that can be pursued. Policy A is business
as usual and Policy B is to take immediate and strong action on climate change. Use these
facts
• If you undertake Policy A there are two possible outcomes. Either warming will occur as
in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or
warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).
Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then
will stabilize at 1.29 degrees and stay that way forever.
• Society is risk neutral
• Use a discount rate of 2%
What is the expected present value of damages up to 2100 under Policy A? What is the
expected present value of damages up to 2100 under Policy B? Suppose undertaking Policy A
costs zero and undertaking Policy B costs X. The total cost of a policy is the implementation
cost plus expected damages. Make a plot of total cost of Policy A and total cost of Policy B
against X. Suppose Congress wants to undertake the policy with the lower total cost. Use your graph to advise Congress.


### **6. Effect of a Risk Averse Society on Advice**
Now suppose that society is risk averse. How do you think this would change your advice to
Congress?


