---
title: "ESM 204 Assignment 4"
author: "Naomi Raal, Virgina Pan, and Cam Audras"
date: "5/22/2022"
output: 
  html_document:
    theme: yeti
    toc: yes
    code_folding: hide # hide code
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(thematic)
library(scales)
library(equatiomatic)
```

```{r}
damage <- read_csv(here("damages.csv"))
warming <- read_csv(here("warming.csv")) %>% 
  clean_names() %>% 
  select(-x1)
```

### **1. Damages vs. Change in Temperature**
Using damages.csv, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature. Estimate a model with a zero intercept because damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data.

```{r}
#create linear regression 
damage$warming_squared <- damage$warming^2
# Solve the issue

quad_damage_model <- lm(damages ~ warming + warming_squared + 0, data = damage)

# define function 
quad_function <- function(warming){
  harm <- quad_damage_model$coefficients[1]*warming + quad_damage_model$coefficients[2]*warming^2
  return(harm)
}
```


```{r}
# visualize function and scatterplot graph
ggplot(quad_damage_model, aes(warming, damages)) + geom_smooth(method = "lm", formula = y ~ x + I(x^2), color = "brown") + geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damage Function", x = "Warming in degrees C", y = "Annual Damages in USD")
```


### **2. Predicted Damages per Year**
Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate and the pulse scenario. Make four plots: (1) damages over time without the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that arises from the pulse, and (4) the difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage).

(1) damages in each year under baseline climate scenario
```{r}
warming$damage_baseline <- quad_function(warming$warming_baseline)
```
```{r}
# visualize function and scatterplot graph
ggplot(data = warming, aes(year, damage_baseline)) + 
  geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damages under baseline", x = "time", y = "Annual Damages in USD")
```

(2) damages in each year under baseline + pulse climate scenario
```{r}
warming$damage_pulse <- quad_function(warming$warming_pulse)
```

```{r}
# visualize function and scatterplot graph
ggplot(data = warming, aes(year, damage_pulse)) + 
  geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damages under baseline with pulse", x = "time", y = "Annual Damages in USD")
```

(3) damages in each year under pulse alone climate scenario
```{r}
warming$damage_pulse_no_baseline <- warming$damage_pulse-warming$damage_baseline
```

```{r}
# visualize function and scatterplot graph
ggplot(data = warming, aes(year, damage_pulse_no_baseline)) + 
  geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damages under pulse alone (no baseline)", x = "time", y = "Annual Damages in USD")
```

(4) difference in damages over time per ton of CO2
```{r}
warming$unit_damage <- (warming$damage_pulse_no_baseline)/(35*10^9)
```

```{r}
# visualize function and scatterplot graph
ggplot(data = warming, aes(year, unit_damage)) + 
  geom_point(size = 0.8) + 
  theme_gray() +
  labs(title = "Damages from pulse per ton of CO2", x = "time", y = "Annual Damages in USD per ton CO2")
```



### **3. Investigating the Relationship between Discount Rate and SCC**
The SCC is the present discounted value of the stream of future damages caused by one
additional ton of CO2. The Obama Administration used a discount rate of 3% to discount
damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot
of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates. Explain the intuition for how the discount rate affects the SCC.

```{r}
# define SCC calculation function 
SCC_calc <- function(principal, r, t, year){
  r = r/100
  SCC <- principal/((1+r)^t)
  return(SCC)
}
```

```{r}
names(warming)[names(warming) == '...1'] <- 'year'
discount_rates_SCC <- warming[, c("year", "unit_damage")]
```

```{r}
discount_rates_SCC$r_1 <- SCC_calc(discount_rates_SCC$unit_damage,1,discount_rates_SCC$year)
total_SCC_1 <- sum(discount_rates_SCC$r_1)
```


```{r}
#Make a for loop to loop through r values
r_vals <- seq(0,10,0.1)
SCC_totals <- vector(mode = "numeric", length = length(r_vals))
count = 1
for(r in r_vals)
{
  
  discount_rates_SCC$r <- SCC_calc(discount_rates_SCC$unit_damage,r,discount_rates_SCC$year)
  SCC_totals[count] <- sum(discount_rates_SCC$r)
  count = count + 1
}
```
```{r}
#plot discount rate vs. SCC
plot(r_vals, SCC_totals, main = "Discount rate vs. SCC", xlab = "discount rate (%)", ylab = "Social Cost of Carbon (SCC) in USD")
```

This plot of discount rate (%) vs. the Social Cost of Carbon (SCC) is consistent with our expectations. We expect the SCC to decrease as the discount rate increases. This is because at a higher discount rate, the value of the future damages is less. Broadly, this means people care less about the future generations than the care about today. Therefore, it is more valuable to emit carbon today and the SCC is low. 

### **4. Using the Ramsey Rule to find SCC**
The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation:

r = ρ + ηg

Using ρ = 0.001, η = 2, and g = 0.01

What is the SCC? Locate this point on your graph from above.

```{r}
#calculate discount rate under Ramsey Rule
rho <-0.001
eta <- 2
g <- 0.01
r_Ramsey <- rho + eta*g
```
```{r}
  discount_rates_SCC$r_Ramsey <- SCC_calc(discount_rates_SCC$unit_damage,(r_Ramsey*100),discount_rates_SCC$year)
  SCC_Ramsey <- sum(discount_rates_SCC$r_Ramsey)
```

Under the Ramsey Rule when discounting within the SCC calculation:

r = ρ + ηg

Using ρ = 0.001, η = 2, and g = 0.01:

r = `r round(r_Ramsey,3)`

SCC = $`r round(SCC_Ramsey,2)` 

Plot with Ramsey discount rate in blue and SCC under Ramsey rule in red. 


```{r}
#plot discount rate vs. SCC
plot(r_vals, SCC_totals, main = "Discount rate vs. SCC", xlab = "discount rate (%)", ylab = "Social Cost of Carbon (SCC) in USD")
abline(v = (r_Ramsey*100), col="blue")
abline(h = SCC_Ramsey, col="red")
```

### **5. Comparing Potential Policy Outcomes**
Now suppose there are two possible climate policies that can be pursued. Policy A is business as usual and Policy B is to take immediate and strong action on climate change. Use these facts:

If you undertake Policy A there are two possible outcomes. Either warming will occur as in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).

Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then
will stabilize at 1.29 degrees and stay that way forever.

Society is risk neutral

Use a discount rate of 2%

What is the expected present value of damages up to 2100 under Policy A? What is the
expected present value of damages up to 2100 under Policy B? Suppose undertaking Policy 
A costs zero and undertaking Policy B costs X. The total cost of a policy is the implementation cost plus expected damages. Make a plot of total cost of Policy A and total cost of Policy B against X. Suppose Congress wants to undertake the policy with the lower total cost. Use your graph to advise Congress.

```{r}
# Estimate PV for the Policy A scenario, discount rate is 2%
discount_rate <- 2
# Need to create a new dataset with the damages and PV for both baseline and 1.5X warming 
damages_pol_a <- warming %>% 
  select(year, warming_baseline, damage_baseline) %>%
  mutate(damages_1.5 = quad_function(1.5*warming_baseline)) %>%
  mutate(pv_baseline =
           SCC_calc(discount_rate, damage_baseline,
                            warming$year, 2022)) %>%
  mutate(pv_1.5 = SCC_calc(discount_rate, damages_1.5,
                            warming$year, 2022))
# Now we need to calculate the probabilities. 
prob_base <- 0.5
prob_1.5 <- 0.5
# Sum the pv for each 
sum_base <- sum(damages_pol_a$pv_baseline)
sum_1.5 <- sum(damages_pol_a$pv_1.5)
# Scale these by probability
expected_value_a <- (sum_base * prob_base) + (sum_1.5 * prob_1.5)
#3023245 billion dollars (I think) CHECK THIS NUMBER 
# Now do this for policy B
damages_policy_b <- warming %>% 
  select(year, warming_baseline) 
for (i in 30:79){
  damages_policy_b$warming_baseline[i]=1.29
}
# Warming locked at 1.29 after 2050
damages_policy_b <- damages_policy_b %>% 
  mutate(damages_b = quad_function(warming_baseline)) %>%
  mutate(pv_b =
           SCC_calc(discount_rate, damages_b,
                            warming$year, 2022))
prob_b <- 1
sum_b <- sum(damages_policy_b$pv_b)  
expected_value_b <- (sum_b * prob_b)
```

- **A: Present Value of Damages under Policy A =** $`r round(expected_value_a, 0)` billion

- **B: Present Value of Damages under Policy B =** $`r round(expected_value_b, 0)` billion

*Assume: Society is risk neutral and the discount rate is 2%. Policy A costs $0 and Policy B costs $X to implement. Total cost is implementation cost + damages.*

#### **Figure 6: Total Cost by Policy Pathway**

```{r}
# Now we need the total cost 
# Cost to implement A is just the expected PV of the damages 
cost_a <- function(x){expected_value_a}
# B adds in the implementation cost (X) 
cost_b <- function(x){expected_value_b + x}
# For the cost of b to equal the cost of a, implementation cost would have to be $2236574 billion: 
x_equal <- expected_value_a - expected_value_b
# Now create a new data frame, including a vector of possible costs (in billions) based on difference calcualted above 
cost <- c(0,3e5,6e5,9e5,12e5,15e5,18e5,21e5,24e5,27e5,30e5)
cost_df <- data.frame(cost) %>% 
  mutate(cost_a = cost_a(cost)) %>% 
  mutate(cost_b = cost_b(cost)) %>% 
  pivot_longer(cols = c(cost_a, cost_b), 
               names_to = 'policy', values_to = 'total_cost') %>%
  mutate(policy = case_when(policy == "cost_a" ~ "Policy A",
                            policy == "cost_b" ~ "Policy B"))
ggplot(data = cost_df, aes(x = cost, y = total_cost)) +
  geom_line(aes(color = policy)) +
  geom_vline(xintercept = x_equal, color = 'darkseagreen', linetype = "dashed") +
  geom_label(aes(x = 2000000, 
                 label = "Implementation Cost Intercept = $2,235,574 billion",
                 y = 1500000), color = "darkseagreen", angle = 90) +
  labs(x = "Implementation Cost ($ billions)", y = "Total Cost ($ billions)")
  
```

**Takeaway:** Congress should adopt **Policy B** unless implementation costs are greater than $`r round(x_equal, 0)` billion. Otherwise, they should adopt Policy A *assuming decisions are made to minimize cost and that total cost of Policy A is the present value of damages through 2100 and total cost of Policy B is the sum of the implementation cost and the present value of those damages through 2100.*

### **6. Effect of a Risk Averse Society on Advice**
Now suppose that society is risk averse. How do you think this would change your advice to Congress?

